---
title: "machinelearning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitness Tracker Analysis and prediction

### Analysis done-
The data was taken and only few columns were chosen for predicting the outcome( accelerometers on the belt, forearm, arm, and dumbell).


#1. Reading the data and selecting only particular columns
```{r }
  library("rpart")
  library("randomForest")
  library("rpart.plot")
  library("dplyr")
  library("reshape2")
  library("caret")
  train<-read.csv("train.csv")
  test<-read.csv("test.csv")

  columns<-grep("*accel*",names(train),value = TRUE)
  data<-select(train,columns)
  data<-select(data,-c(var_total_accel_belt,var_accel_arm,var_accel_dumbbell,var_accel_forearm))
  
  classe<-train$classe
  
  
  data<-cbind(data,classe=classe)
```


2.Plotting all the varibales against classe

```{r}
  df.m <- melt(data, "classe")
  ggplot(df.m,aes(x=value,y=classe))+geom_point()+facet_wrap(~variable)
```


3. Trying different method for training-
```{r}
  ##setting number=5 dues to memory constraints
  tr<-trainControl(method = "cv", number = 5)
  cartGrid = expand.grid( .cp = seq(0.00002,0.002,0.00002))
  model<-train(classe~.,data=data,method="rpart",trControl=tr,tuneGrid=cartGrid)
  ##accuracy of 82% at cp=0.0002
  prp(model$finalModel)
  getTrainPerf(model)
  
  
  ##randomForest accuracy ~94%
  model2<-randomForest(classe~.,data=data)
  predict(model2,newdata = test)
  ##getTrainPerf(model2)
  
  
  model3<-train(classe~.,data=data,method="rf",trControl=trainControl(method="cv",number=5))
  ##accuracy of ~95% with cv
  getTrainPerf(model3)
```

4. Selecting model with highest accuracy and predicting the test data with it-
```{r}
  predictionResults<-predict(model3,newdata = test)
  print(predictionResults)
```